---
title: "photoclean"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{photoclean}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(svdtools)
```

Single Value Decomposition can remove noise from matrices in the process of reducing them. One example is image filtering. The included data object _noisymatrix_ is one such example. This is a grayscale image of Linux Penguin on Wikimedia Commons from Larry Ewing <lewing@isc.tamu.edu>. At random, some pixels have been deleted, creating noise.
```{r, echo=FALSE}
image(noisymatrix, col=gray.colors(65536))
```

There are several ways we could filter this image to try to remove the impact of the bad pixels. Let's examine the explanation plots.

```{r}
plot_explanation(noisymatrix, limit=.99)
```

There are a lot of components to play with. One brute force way of doing this would be to try specifying some percentage of components to keep. We could try 98%:
```{r}
m<-reduce_percentage(noisymatrix, .98)
image(m, col=gray.colors(65536))
```
This image is obviously overfiltered. Instead, we could simply eliminate all the components above some threshold - for example, discard 51-411:
```{r}
# Keep the first 50 components
m<-reduce_components(noisymatrix, 50)
image(m, col=gray.colors(65536))
```

This looks significantly better, but we could try tuning the image to keep more of the data, and removing only a block of components. We can try excluding 51-250:
```{r}
m<-exclude_components(noisymatrix,51:250)
image(m, col=gray.colors(65536))

```

In this way, we can gradually select between the level of sharpness and the level of noise we retain!
